Copulas and their potential for ecology

### 2019-10-07 Paper.Rmd
finished an edit of Discussion, made a list of what else needs to be done

# Abstract
<!--***DANr: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
All branches of ecology study relationships among environmental and biological variables. However, 
standard approaches to studying such relationships, based on correlation and regression, 
provide only a small slice of the complex information contained in the relationships. Other 
statistical approaches exist that provide a complete description of relationships between variables, 
based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
but not in ecology. We here explore the concepts that underpin copulas and examine the potential 
for those concepts to improve our understanding of ecology. We find that informative copula 
structure in dependencies between variables is exceedingly common across all the environmental, 
species-trait, phenological, population, community, and ecosystem functioning datasets we 
considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
effect whereby dependence structures between environmental variables are inherited by ecological 
variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
under Liebig's law of the minimum. Copula structure and tail 
dependence can also help reveal causal relationships between variables. We also describe consequences of copula structure for ecological phenomena, including impacts 
on extinction risk and the stability through time of ecosystem services. By documenting the 
importance of a complete description of dependence, advancing conceptual frameworks, and 
demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology, benefiting the discipline.

@@

Thus Sklar's theorem implies that
any random vector can be constructed from a copula and marginal 
distributions. And conversely,
any copula and any univariate distribution functions give rise to a random vector. 
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables. Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. For instance,
construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
in the left than in the right tails. Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].
<!--***DAN: dont forget to put these few words in!-->

@@

<!--***DAN: In the Discussion, you need to point out not only that we
use only bivariate copulas (and there are multivariate ones), but also
that we use only continuous marginals. Probably an adequate approximation
for our data, but see the other papers I've seen recently which use count data,
and discrete marginals. Approaches such as that will be necessary for some 
datasets-->

@@

<!--***DAN: Note: I am using tail dependence for the specific thing defined on p 62 of Joe,
and tail associations as a more general term. I have changed this in parts of the ms
upsteam from this point, but need to remember to change it as I move downstream.-->

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
<!--***DAN: dont forget to put these in!-->
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distribution of these values across location pairs.

@@

For each $b$ and $g_i$, we simulated the model for $25000$ time
steps and retained the $E_i(t)$ for the final $2500$ time 
steps. We applied our 
nonparametric statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
and our Spearman- or Kendall-preserving normal-copula 
surrogate comparison methods (see section on methods for Q1) to these 
outputs to discover if the model could
produce asymmetric tail associations (and therefore non-normal copula structure)
between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
from bivariate normal distributions), the Moran 
mechanism analyzed below does not operate here.
<!--***DAN: You may also want to have some additional commentary in the Discussion on
how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
want to insert a point at the end of the Methids section here again emphasizing
that these are simple models for an initial exploration, and we discuss in the 
Discussion how...-->

@@

<!--results for third mechanism-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/BMR_results/evo_cop/singlepage_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
measures of asymmetry of tail associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ K-O) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
(E, J, O)). See the methods section for Q2 for details. The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}

@@

To explore the influence of copula structure and tail associations on spatial Taylor's
law, we carried out a series of simulations that generated sets of population 
matrices that had 
distinct copula structure between locations but that 
were otherwise statistically similar. For $k=1,\ldots,1000$
and X representing a Clayton, normal, Frank, or a survival Clayton copula family,
we generated $N \times n$ population matrices $m^{(\text{X},k)}$ with the following properties:
1) for any given values of $j$ and $k$, the values 
$\{ m_{t,j}^{(\text{X},k)} : t=1,\ldots,N  \}$ were exactly the same, as unordered
sets, for all X, i.e., the same actual population values were used for location $j$ 
in simulation $k$, regardless of X;
2) for any given $k$, $j_1 \neq j_2$, the Spearman correlations (computed through time)
$\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
variation, i.e., correlations through time of populations in two locations were the same,
up to sampling variation, regardless of X;
3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was X.
We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
for X taking the values Clayton, normal, Frank, and survival Clayton
were statistically similar except for the copula structure between locations, which was
X. Thus comparing how Taylor's law may manifest differently for $m^{(\text{X},k)}$
for different values of X constitutes a test of the influence of copula
structure on Taylor's law. We selected X equal to the Clayton, normal, Frank, and survival 
Clayton copulas to explore a range of tail association patterns.

A variety of Taylor's law statistics were computed for each populaton matrix
$m^{(\text{X},k)}$. First, the $N=50$ 
spatial means,
$m_g$, and variances, $v_g$, were computed, and $\log(v_g)$ vs. $\log(m_g)$ plots 
were considered. Linearity of the $\log(v_g)$ vs. $\log(m_g)$ relationship was 
tested for each simulation by comparing the linear regression
through these $50$ points to a quadratic alternative via an $F$-test, 
producing a $p$-value result. If 
these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
simulations which were generated for X, it supported the 
linearity assumption of Taylor's law for X; whereas if 
they were clustered toward smaller values
the test tended to reject that assumption. We also tested, for each simulation, 
the assumption of Taylor's 
law that the $\log(v_g)$ vs. $\log(m_g)$ plot was homoskedastic: we regressed the absolute 
residuals of the 
linear regression of $\log(v_g)$ versus $\log(m_g)$ against the predictions of that 
regression. A significant $p$-value
result of this test indicates heteroskedasticity. For each simulation we also computed the
root mean squared error of $\log(v_g)$ vs. $\log(m_g)$ data from the linear regression
of $\log(v_g)$ versus $\log(m_g)$, as well as the 
intercept and slope of the linear regression.
Finally we recorded the quadratic coefficient of the regression of $\log(v_g)$ against
$\log(m_g)$ and $(\log(m_g))^2$, and the mean curvature of 
the quadratic regression equation itself
across the values $\log(m_g)$. The distrubutions of all these 
statistics across the 1000 replicate 
simulations were compared for X the Clayton, normal, Frank, and survival Clayton copulas, 
to determine if copula structure influences Taylor's law.

@@

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.8 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in red and green to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See the methods
section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}
<!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
matched up with the spacing of the panels themselves, can you fix it please?
***Shya: check now, if is it ok?-->

@@

<!--what we did and why is it important to study?-->
We showed that non-normal copula structure and asymmetric tail dependence are 
common across multiple sub-disciplines
in ecology, although no research we know of has 
taken this structure into account. We hypothesized 
mechanisms that may cause non-normal copula structures and tail dependence;
we discuss below how commonly some of our
mechanisms may opperate. 
We also hypothesized and demonstrated important consequences of 
non-normal copula structure and tail-dependence for the field of ecology. 
For instance, the skewness
of a spatial-average time series is influenced by tail dependencies between its 
constituent time series: right-tail-dependent local time series 
lead to "spiky" spatially averaged time series, with large outbreaks.
Thus tail dependence could have implications for exploited and pest species.
Extinction risk can also be altered by tail dependence across space
in environmental variables. 
In our view, our results make it reasonable to suggest 
that comprehensive understanding 
of many ecological phenomena may be impossible without 
a complete copula characterization of dependence.

Copula approaches are
well developed [@joe2014_dependence; @nelsen2006_copula], and 
computer implementations exist (`copula` and 
`VineCopula` packages). Ecologists can apply these tools
immediately. We created several interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that build upon existing copula methods. 
To our knowledge, only normal copulas have 
been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
Such use is equivalent to using normal distributions with transformed
marginals; i.e., it is equivalent to an approach not using copulas at all, and
therefore does not capitalize on the power of copula techniques.

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first causal mechanism (Fig. \ref{MasterFigure}, A) probably operates commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are so common, it is reasonable to suppose the mechanism
opperates commonly. It may be a dominant cause of 
tail depencence and non-normal copula struture of ecological dependencies 
across space.
We provide some empirical support for the mechanism 
in our discussion of green sprice aphids and
winter temperature below.

<!--para on how likely one is to get copula structure via the Moran mechanism-->
There are also reasons to expect that 
our Moran mechanism (Fig. \ref{MasterFigure}, B) 
operates commonly: Moran effects are common [@Liebhold2004], and 
non-normal and tail-dependent copula structures are often found in 
environmental variables. If intense meteorological events are 
also widespread, then
environmental variables associated with these events should take extreme
values simultaneously across large spatial areas, producing tail dependence in measurements
made through time at different locations. Non-extreme values may instead
be associated with local phenomena, and therefore may be
less correlated across large areas.
@Serinaldi2008 examined the spatial dependence 
of rainfall in Central Italy; Gumbel or Student 2-copulas 
were candidates for modelling dependence, and 
neither of these is a normal copula.
A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula effectively modeled 
the spatial dependence of drought variables.
Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas using spatial datasets [@Goswami2018].

<!--Causal mechanism C, which we also did not test-->
We suggested in the Introduction that asymmetric competitive relationships
between species could yield tail dependence between abundance data for the 
species. This is another theoretical mechanism for non-normal copula
structure and tail dependence, represented in Fig. \ref{MasterFigure} 
by the box "Asymmetric species interactions" and the arrow labelled "C".
It could be tested by analyzing copulas of abundances of competing 
species, sampled across space or time.

<!--Holder stuff-->
Our simulations of character evolution suggest the hypothesis that
changes through evolutionary time in bird and 
mammal BMR and body size may exhibit right-tail dependence,
contrary to standard normality assumptions of character evolution models. This is
a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
could have come about in another, unknown way. Our simulations show that tail dependence 
in evolutionary changes is sufficient, but may not be necessary, to produce tail
dependence in characters of extant species. For instance (see below, 
Appendix \ref{SM-sect:missingdata}, Fig. \ref{SM-fig:missingdata}), 
systematically missing data can also produce tail dependence and may
have influenced results for the BMR-body mass datasets. Even if the hypothesized 
evolutionary mechanism (Fig. \ref{MasterFigure}, D)
is correct, our results only replace one question, i.e., why do we see right-tail
dependence between BMR and body mass, with another,
i.e., why might we see right-tail dependence in evolutionary changes in these traits?
See Appendix \ref{SM-evol_mech} for further thoughts on potential importance of 
copulas for character evolution.

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
Relationships between BMR and body mass relate
to a trade-off between mass-specific BMR (BMR per unit body mass) and body
mass itself. Copulas probably interrelate with 
life-history trade-off theory in additional ways beyond what we demonstrated.
For instance, it is well known that energy allocation to a life function, F (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $G_1$, $G_2$, $G_3$ 
(e.g., growth, predation avoidance). This is the principle of 
allocation. But F can trade off against any or all of 
the $G_i$. Therefore, for large F, approaching absolute 
limitations, there may be a strong 
association between F and $G_1$. For small F, there may be
little association because resources not allocated to F can instead be allocated 
to any combination of the $G_i$. This constitutes 
asymmetric tail dependence between F and $G_1$. @Winemiller1992
described a three-way trade-off in fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little such association for later-maturing
fishes because those fishes may invest the resources not invested in maturing quickly into
either fecundity or juvenile survival.
These ideas suggest that copulas will be 
useful for studying multi-dimensional trade-offs. But applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the characters, then statistical artefacts
can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}).

>>>>

<!--***DAN: remianing to do:
1) Edit abstract
2) Go through the whole thing one more time
3) send it out to coauthors
4)make sure to say something about tropical systems
5) maybe write to the editors saying this is coming soon?
6) In the Discussion, you maybe need to point out not only that we
use only bivariate copulas (and there are multivariate ones), but also
that we use only continuous marginals. Probably an adequate approximation
for our data, but see the other papers I've seen recently which use count data,
and discrete marginals. Approaches such as that will be necessary for some 
datasets
7) change the box around the Cedar Creek data in the main fig
-->

# Abstract
<!--***DAN: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
All branches of ecology study relationships among environmental and biological variables. However, 
standard approaches to studying such relationships, based on correlation and regression, 
provide only a small slice of the complex information contained in the relationships. Other 
statistical approaches exist that provide a complete description of relationships between variables, 
based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
but not in ecology. We here explore the concepts that underpin copulas and examine the potential 
for those concepts to improve our understanding of ecology. We find that informative copula 
structure in dependencies between variables is exceedingly common across all the environmental, 
species-trait, phenological, population, community, and ecosystem functioning datasets we 
considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
effect whereby dependence structures between environmental variables are inherited by ecological 
variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
under Liebig's law of the minimum. Copula structure and tail 
dependence can also help reveal causal relationships between variables. We also describe consequences of copula structure for ecological phenomena, including impacts 
on extinction risk and the stability through time of ecosystem services. By documenting the 
importance of a complete description of dependence, advancing conceptual frameworks, and 
demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology, benefiting the discipline.

@@

Thus Sklar's theorem implies that
any random vector can be constructed from a copula and marginal 
distributions. And conversely,
any copula and any univariate distribution functions give rise to a random vector. 
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables. Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. For instance,
construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
in the left than in the right tails. Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].

@@

@@

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distribution of these values across location pairs.

@@

For each $b$ and $g_i$, we simulated the model for $25000$ time
steps and retained the $E_i(t)$ for the final $2500$ time 
steps. We applied our 
nonparametric statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
and our Spearman- or Kendall-preserving normal-copula 
surrogate comparison methods (see section on methods for Q1) to these 
outputs to discover if the model could
produce asymmetric tail associations (and therefore non-normal copula structure)
between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
from bivariate normal distributions), the Moran 
mechanism analyzed below does not operate here.

@@

<!--results for third mechanism-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/BMR_results/evo_cop/singlepage_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in 
character distributions]{Three measures of asymmetry of tail 
associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$, K-O) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
(E, J, O)). See the methods section for Q2 for details. The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}

@@

To explore the influence of copula structure and tail associations on spatial Taylor's
law, we carried out a series of simulations that generated sets of population 
matrices that had 
distinct copula structure between locations but that 
were otherwise statistically similar. For $k=1,\ldots,1000$
and X representing a Clayton, normal, Frank, or a survival Clayton copula family,
we generated $N \times n$ population matrices $m^{(\text{X},k)}$ with the following properties:
1) for any given values of $j$ and $k$, the values 
$\{ m_{t,j}^{(\text{X},k)} : t=1,\ldots,N  \}$ were exactly the same, as unordered
sets, for all X, i.e., the same actual population values were used for location $j$ 
in simulation $k$, regardless of X;
2) for any given $k$ and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the Spearman correlations (computed through time)
$\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
variation, i.e., correlations through time of populations in two locations were the same,
up to sampling variation, regardless of X;
3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was X.
We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
for X taking the values Clayton, normal, Frank, and survival Clayton
were statistically similar except for the copula structure between locations, which was
X. Thus comparing how Taylor's law may manifest differently for $m^{(\text{X},k)}$
for different values of X constitutes a test of the influence of copula
structure on Taylor's law. We selected X equal to the Clayton, normal, Frank, and survival 
Clayton copulas to explore a range of tail association patterns.

A variety of Taylor's law statistics were computed for each populaton matrix
$m^{(\text{X},k)}$. First, the $N=50$ 
spatial means,
$m_g$, and variances, $v_g$, were computed, and $\log(v_g)$ vs. $\log(m_g)$ plots 
were considered. Linearity of the $\log(v_g)$ vs. $\log(m_g)$ relationship was 
tested for each simulation by comparing the linear regression
through these $50$ points to a quadratic alternative via an $F$-test, 
producing a $p$-value result. If 
these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
simulations which were generated for X, it supported the 
linearity assumption of Taylor's law for X; whereas if 
they were clustered toward smaller values
the test tended to reject that assumption. We also tested, for each simulation, 
the assumption of Taylor's 
law that the $\log(v_g)$ vs. $\log(m_g)$ plot was homoskedastic: we regressed the absolute 
residuals of the 
linear regression of $\log(v_g)$ versus $\log(m_g)$ against the predictions of that 
regression. A significant $p$-value
result of this test indicates heteroskedasticity. For each simulation we also computed the
root mean squared error of $\log(v_g)$ vs. $\log(m_g)$ data from the linear regression
of $\log(v_g)$ versus $\log(m_g)$, as well as the 
intercept and slope of the linear regression.
Finally we recorded the quadratic coefficient of the regression of $\log(v_g)$ against
$\log(m_g)$ and $(\log(m_g))^2$, and the mean curvature of 
the quadratic regression equation 
across the values $\log(m_g)$. The distrubutions of all these 
statistics across the 1000 replicate 
simulations were compared for X the Clayton, normal, Frank, and survival Clayton copulas, 
to determine if copula structure influences Taylor's law.

@@

<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
\begin{figure}[!h]
\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
\vspace{-0.8 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in red and green to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were similar. See the methods
section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
\end{center}
\end{figure}

@@

<!--what we did and why is it important to study?-->
We showed that non-normal copula structure and asymmetric tail associations are 
common across multiple sub-disciplines
in ecology, although these facets of ecological data are only ocassionally 
taken into account [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
mechanisms that may cause non-normal copula structures and asymmetric tail associations;
we discuss below how commonly some of our
mechanisms may opperate. 
We also hypothesized and demonstrated important consequences of 
non-normal copula structure and asymmetric tail associations for the field of ecology. 
For instance, the skewness
of a spatial-average time series is influenced by asymmetric tail associations between its 
constituent time series: predominantly right-tail-associated local time series 
can lead to "spiky" spatially averaged time series, with large outbreaks.
Thus tail associations could have implications for exploited and pest species.
Extinction risk and Taylor's law can also be altered by tail association patterns across space
in environmental variables. 
In our view, our results make it reasonable to suggest 
that a more comprehensive understanding 
of many ecological phenomena may be possible if a
complete copula characterization of dependence is employed. 
Copula approaches are
well developed [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], statistically, 
and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
open-source computer implementations exist (e.g., the `copula` and 
`VineCopula` packages in the R programming language). Ecologists can apply these tools
immediately. We created several interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that build upon existing copula methods.

<!--Stuff about the mechanisms-->

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are so common, it is reasonable to hypothesize that the mechanism
may opperate commonly. It may be a dominant cause of 
asymmetric tail associations and non-normal copula struture of ecological 
dependencies across space.
We provide some further support for the mechanism 
in our discussion of green spruce aphids and
winter temperature below.

<!--para on how likely one is to get copula structure via the Moran mechanism-->
There are also reasons to hypothesize that 
our Moran mechanism (Fig. \ref{MasterFigure}, B) 
may operate commonly: Moran effects are common [@Liebhold2004], and 
non-normal copula structures and asymmetric tail associations are often found in 
environmental variables. If intense meteorological events are 
also widespread, then
environmental variables associated with these events should take extreme
values simultaneously across large spatial areas, producing tail dependence in measurements
made through time at different locations. Non-extreme values may instead
be associated with local phenomena, and therefore may be
less correlated across large areas.
@Serinaldi2008 examined the spatial dependence 
of rainfall in Central Italy; Gumbel or Student 2-copulas 
were candidates for modelling dependence, and 
neither of these is a normal copula.
A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula effectively modeled 
the spatial dependence of drought variables; the Gumbel copula demonstrates 
asymmetric tail associations.
Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas using spatial datasets [@Goswami2018].

<!--Causal mechanism C, which we also did not test-->
We suggested in the Introduction that asymmetric competitive relationships
between species could yield asymmetric tail associations between abundance measurements for the 
species. This is another theoretical mechanism for non-normal copula
structure and asymmetric tail associations, represented in Fig. \ref{MasterFigure} 
by the box "Asymmetric species interactions" and the arrow labelled "C".
It could be tested by analyzing copulas of abundances of competing 
species, sampled across space or time.

<!--Holder stuff-->
Our simulations of character evolution suggest the hypothesis that
changes through evolutionary time in bird and 
mammal BMR and body size may exhibit greater right- than left-tail association,
contrary to standard normality assumptions of character evolution models. This is
a hypothesis only, because the greater right-tail association shown in 
Fig. \ref{fig_biv_multi_raw_cop}F,G
could have come about in another, unknown way rather than via the mechanism we suggested
which implicated asymmetric tail associations in evolutionary change. 
Our simulations show that asymmetric tail associations 
in evolutionary changes are sufficient, but may not be necessary, to produce 
the observed asymmetric tail
associations in characters of extant species. For instance (see below, and Appendix
\ref{SM-sect:missingdata} and Fig. \ref{SM-fig:missingdata}), 
systematically missing data can also produce tail dependence and may
have influenced results for the BMR-body mass datasets. Even if the hypothesized 
evolutionary mechanism (Fig. \ref{MasterFigure}, D)
is correct, our results only replace one question, i.e., why do we see greater right-
than left-tail
association between BMR and body mass, with another,
i.e., why might we see greater right- than left-tail association in evolutionary 
changes in these traits?

Performing statistical tests for associations between continuous traits in different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    [@Felsenstein1985] is currently the second most-cited paper in the history
    of the journal the American Naturalist [@HueyGT2019].
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which 
    the duration of the branch in time can be treated as a proxy for the branch length.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study [@Ives2018;@Harmon2018].
Even if one were able to simply use a time-based set of branch lengths, assigning
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.
